Reproducible Research: Peer Assessment 1
========================================================



## Loading and preprocessing the data



```{r}
data <- read.csv("activity.csv")
summary(data)
```
Transform the date variable into Date format
```{r}
data$Date<-as.Date(data$date)
```




## What is mean total number of steps taken per day?




Calculate the total number of steps taken each day
```{r}
steps.per.day<-tapply(data$steps, data$Date, FUN=sum)
```

Histogram of the total number of steps taken each day
```{r fig.width=7, fig.height=6}
hist(steps.per.day, breaks =15, main = "Histogram of Steps Taken per Day", xlab="total steps per day", col="salmon")
```

Calculate and report the mean and median total number of steps taken per day
```{r}
steps.mean<-mean(steps.per.day, na.rm=TRUE)
steps.median<-median(steps.per.day, na.rm=TRUE)
```
The mean total number of steps taken per day is `r format(steps.mean, digits=7)`   
The median total number of steps taken per day is `r steps.median`


## What is the average daily activity pattern?


Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r fig.width=7, fig.height=6}
plot(y=data$steps, x=data$Date, type="l", col="blue", main="Average number of steps taken by day", ylab="number of steps")
```

  
## Imputing missing values

Calculate and report the total number of missing values in the dataset
```{r}
NAs<-sum(is.na(data$steps))
```
The total number of missing values in the dataset is `r NAs`.

I use the mean value of steps taken every five minutes to fill in the missing values. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
data2=data
row_NAs <- which(is.na(data2$steps))
data2$steps<-replace(data2$steps, row_NAs, mean(data$steps, na.rm=TRUE))
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
steps.per.day2<-tapply(data2$steps, data$Date, FUN=sum)
```

Histogram of the total number of steps taken each day
```{r fig.width=7, fig.height=6}
hist(steps.per.day2, breaks=15, main = "Histogram of Steps Taken per Day \n(missing value filled)", xlab="total steps per day", col="salmon")
```

The two graphs are different. The second graph where the missing values are filled have higher frequencies around the mean 10766.19


  
## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
data2$weekdays<-weekdays(data2$Date)
data2$weekdays2<-ifelse(data2$weekdays %in% c("Saturday", "Sunday"), "weekend", "weekday") 
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r fig.width=7, fig.height=8}
library(lattice) 
xyplot(steps ~ interval | weekdays2, data = data2, type = "l", ylab="Number of steps")
```

Yes there are differences in actvity patterns between weekdays and weekends.
