Reproducible Research: Peer Assessment 1
========================================================

setwd("D:/Users/xu/Desktop/R Training/Coursera/repdata_data_activity")

## Loading and preprocessing the data

```{r}
data <- read.csv("activity.csv")
summary(data)
```

Transform the date into Date format
```{r}
data$Date<-as.Date(data$date)
```

## What is mean total number of steps taken per day?

Calculate the total number of steps taken each day
```{r}
steps.per.day<-tapply(data$steps, data$Date, FUN=sum)
```

Histogram of the total number of steps taken each day
```{r fig.width=7, fig.height=6}
hist(steps.per.day, breaks =15, main = "Histogram of Steps Taken per Day", xlab="total steps per day", col="salmon")
```

Calculate and report the mean and median total number of steps taken per day
```{r}
steps.mean<-mean(steps.per.day, na.rm=TRUE)
steps.median<-median(steps.per.day, na.rm=TRUE)
```
The mean total number of steps taken per day is `r steps.mean`  
The median total number of steps taken per day is `r steps.median`


## What is the average daily activity pattern?

Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r fig.width=7, fig.height=6}
plot(y=data$steps, x=data$Date, type="l", col="blue", main="Average number of steps taken by day", ylab="number of steps")
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset
```{r}
NAs<-sum(is.na(data$steps))
```
The total number of missing values in the dataset is `r NAs`.

I use the mean value of steps taken every five minutes to fill in the missing values. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
data2=data
row_NAs <- which(is.na(data2$steps))
data2$steps<-replace(data2$steps, row_NAs, mean(data$steps, na.rm=TRUE))
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
steps.per.day2<-tapply(data2$steps, data$Date, FUN=sum)
```
Histogram of the total number of steps taken each day
```{r fig.width=7, fig.height=6}
hist(steps.per.day2, breaks=15)
```
The two graphs are different.

## Are there differences in activity patterns between weekdays and weekends?



